---
title: "Data 201 Deliverable 4"
author: "Anibal"
date: "2026-09-15"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
library(tidyverse)

files <- list.files(
  path = "data/raw",
  pattern = "\\.csv$",
  full.names = TRUE)


airbnb_data <- map_dfr(files, function(file) {
 data <- read_csv(file, show_col_types = FALSE)
  data$month_year <- tools::file_path_sans_ext(basename(file))

  data
})
```

```{r}
nrow(airbnb_data)
```
```{r}
names(airbnb_data)
```
```{r}
christchurch_data <- airbnb_data %>%
  filter(neighbourhood_group == "Christchurch City")
```

```{r}
table(christchurch_data$month_year)
```
```{r}
colSums(is.na(christchurch_data))
```
```{r}
sum(duplicated(christchurch_data))
```
```{r}
summary(christchurch_data)
```
```{r}
christchurch_data %>%
  summarise(
    missing_price = sum(is.na(price)),
    missing_minimum_nights = sum(is.na(minimum_nights)),
    missing_last_review = sum(is.na(last_review)),
    missing_reviews_per_month = sum(is.na(reviews_per_month)),
    missing_license = sum(is.na(license))
  )
```
```{r}
summary(christchurch_data$price)
```
```{r}
table(christchurch_data$room_type)
```
```{r}
christchurch_data %>%
  filter(price > 1000) %>%
  select(id, name, price, room_type, neighbourhood) %>%
  arrange(desc(price))
```
```{r}
christchurch_data %>%
  filter(price <= 0) %>%
  select(id, name, price, room_type)
```
```{r}
quantile(
  christchurch_data$price,
  probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1),
  na.rm = TRUE
)
```
```{r}
christchurch_data %>%
  filter(price > 1000) %>%
  count(price, sort = TRUE)
```
```{r}
christchurch_data %>%
  filter(price > 1000) %>%
  summarise(
    number_of_listings = n(),
    lowest_high_price = min(price, na.rm = TRUE),
    highest_price = max(price, na.rm = TRUE),
    median_high_price = median(price, na.rm = TRUE)
  )
```
```{r}
christchurch_data %>%
  filter(price > 1000) %>%
  count(price, sort = TRUE) %>%
  slice_head(n = 15)
```
```{r}
christchurch_clean <- christchurch_data %>%
  filter(!is.na(price), price <= 10000)
```

```{r}
christchurch_clean <- christchurch_clean %>%
  select(-license)
```

```{r}
christchurch_clean %>%
  summarise(
    invalid_minimum_nights = sum(minimum_nights <= 0, na.rm = TRUE),
    invalid_availability = sum(availability_365 < 0 | availability_365 > 365, na.rm = TRUE),
    invalid_latitude = sum(latitude < -90 | latitude > 90, na.rm = TRUE),
    invalid_longitude = sum(longitude < -180 | longitude > 180, na.rm = TRUE),
    invalid_reviews = sum(number_of_reviews < 0, na.rm = TRUE),
    invalid_review_rate = sum(reviews_per_month < 0, na.rm = TRUE)
  )
```
```{r}
christchurch_clean <- christchurch_clean %>%
  filter(!is.na(minimum_nights))
```
```{r}
colSums(is.na(christchurch_clean))
```
```{r}
dim(christchurch_clean)
```

```{r}
write_csv(
  christchurch_clean,
  "data/clean/christchurch_clean.csv"
)
```

Airbnb Data Cleaning

The nine monthly Airbnb datasets from October 2025 to June 2026 were combined into one dataset. The data was then filtered to only include listings from Christchurch City. A `month_year` column was added so the listings could be compared across the different months. The original Christchurch dataset contained 28,795 rows and 19 columns.

The data was then cleaned by removing listings with missing prices, as price is needed to compare rental costs. Listings with prices above $10,000 were also removed because some of the extreme prices did not match the actual prices found online and appeared to be data errors. The $10,000 cutoff was chosen to remove obvious errors while still keeping potentially legitimate high-priced properties. Rows with missing `minimum_nights` were removed because this variable may be useful for future analysis.

The `license` column was removed because every value in this column was missing. No duplicate rows were found. Missing values in `last_review` and `reviews_per_month` were kept because these may represent listings that have not received any reviews. Removing these listings would reduce the dataset unnecessarily. Latitude and longitude were kept because they may be useful for future location-based analysis.

After cleaning, the dataset contained 18,110 rows and 18 columns. The cleaned dataset was saved as `christchurch_clean.csv` in the `data/cleaned` folder.


## Tenancy Services Rental Bond Data


```{r tenancy-import}

library(tidyverse)

tenancy_data <- read_csv(
  "data/tenancy_raw/Detailed-Quarterly-Tenancy-Q1-2020-Q3-2026.csv",
  show_col_types = FALSE
)
```

```{r}
dim(tenancy_data)
```
```{r tenancy-columns}
names(tenancy_data)
```
```{r}
unique(tenancy_data$TimeFrame)
```
```{r}
tenancy_data <- tenancy_data %>%
  mutate(TimeFrame = as.Date(TimeFrame))
```

```{r}
sort(unique(tenancy_data$TimeFrame))
```
```{r}
tenancy_filtered <- tenancy_data %>%
  filter(
    TimeFrame >= as.Date("2025-10-01"),
    TimeFrame <= as.Date("2026-04-01")
  )
```

```{r}
dim(tenancy_filtered)
```
```{r}
sort(unique(tenancy_filtered$TimeFrame))
```
```{r}
head(tenancy_filtered)
```
```{r}
glimpse(tenancy_filtered)
```

```{r}
tenancy_clean <- tenancy_filtered %>%
  mutate(
    across(
      c(
        `Median Rent`,
        `Geometric Mean Rent`,
        `Upper Quartile Rent`,
        `Lower Quartile Rent`,
        `Log Std Dev Weekly Rent`
      ),
      ~ na_if(.x, "NULL")
    ),
    `Number Of Beds` = na_if(`Number Of Beds`, "NULL"),
    `Location Id` = na_if(`Location Id`, "NULL")
  )
```

```{r}
unique(tenancy_filtered$`Number Of Beds`)
```
```{r}
sum(is.na(tenancy_clean$`Number Of Beds`))
```
```{r}
tenancy_clean <- tenancy_filtered %>%
  mutate(
    across(
      c(
        `Median Rent`,
        `Geometric Mean Rent`,
        `Upper Quartile Rent`,
        `Lower Quartile Rent`,
        `Log Std Dev Weekly Rent`
      ),
      ~ na_if(.x, "NULL")
    ),
    `Number Of Beds` = na_if(`Number Of Beds`, "NULL"),
    `Location Id` = na_if(`Location Id`, "NULL")
  ) %>%
  mutate(
    `Median Rent` = as.numeric(`Median Rent`),
    `Geometric Mean Rent` = as.numeric(`Geometric Mean Rent`),
    `Upper Quartile Rent` = as.numeric(`Upper Quartile Rent`),
    `Lower Quartile Rent` = as.numeric(`Lower Quartile Rent`),
    `Log Std Dev Weekly Rent` = as.numeric(`Log Std Dev Weekly Rent`),
    `Number Of Beds` = factor(
      `Number Of Beds`,
      levels = c("0", "1", "2", "3", "4", "5", "5+", "6", "7", "9", "ALL")
    )
  )
```

```{r}
glimpse(tenancy_clean)
```
```{r}
colSums(is.na(tenancy_clean))
```
```{r}
unique(tenancy_clean$`Location Id`)[1:30]
```

```{r}
tenancy_clean %>%
  count(`Location Id`, sort = TRUE) %>%
  head(15)
```
```{r}
tenancy_clean %>%
  filter(is.na(`Location Id`) | `Location Id` == "-99") %>%
  select(TimeFrame, `Location Id`, `Dwelling Type`, `Number Of Beds`,
         `Total Bonds`, `Median Rent`) %>%
  head(20)
```
```{r}
tenancy_clean %>%
  filter(is.na(`Location Id`) | `Location Id` == "-99") %>%
  count(`Location Id`, `Number Of Beds`, `Dwelling Type`)
```
```{r}
tenancy_clean %>%
  filter(is.na(`Location Id`)) %>%
  select(
    TimeFrame,
    `Location Id`,
    `Dwelling Type`,
    `Number Of Beds`,
    `Total Bonds`,
    `Median Rent`
  ) %>%
  head(20)
```
```{r}
tenancy_clean <- tenancy_clean %>%
  filter(!is.na(`Location Id`))
```

```{r}
dim(tenancy_clean)
```
Rows with missing location IDs were removed because they had no median rent values. Rows with a location ID of -99 were retained because they contained valid bond and rent information.

```{r}
colSums(is.na(tenancy_clean))
```
```{r}
tenancy_clean %>%
  filter(is.na(`Number Of Beds`)) %>%
  select(
    TimeFrame,
    `Location Id`,
    `Dwelling Type`,
    `Number Of Beds`,
    `Total Bonds`,
    `Median Rent`
  ) %>%
  head(20)
```
Missing values in Number Of Beds were retained because the rows still contained valid location, dwelling type, bond, and rent information. Removing them would result in a loss of useful rental data.

```{r}
write_csv(
  tenancy_clean,
  "data/Clean/tenancy_clean.csv"
)
```

```{r}
list.files("data/Clean")
```


### Tenancy Data Cleaning

The tenancy dataset was filtered to match the Airbnb timeframe from October 2025 to April 2026. `"NULL"` values were converted to missing values, and the rent columns were changed to numeric values for analysis. `Number Of Beds` was kept as a categorical variable because it included values such as `"ALL"` and `"5+"`. Rows with missing `Location Id` values were removed because they did not contain useful rent information. However, rows with `Location Id` values of `"-99"` and missing bedroom information were kept because they still contained useful rental and bond data. The final cleaned dataset contained 27,118 rows.



