---
title: "Data Wrangling - Python Workflow Converted to R"
author: "Sohiththarun"
output: html_document
---

## 1. Load and inspect the October 2025 dataset

```{r}
library(tidyverse)
library(lubridate)

df <- read_csv("October_2025.csv", show_col_types = FALSE)

head(df)
```

```{r}
head(df)
```

```{r}
dim(df)
```

```{r}
glimpse(df)
```

## 2. Filter Christchurch only

```{r}
christchurch <- df %>%
  filter(neighbourhood_group == "Christchurch City")
```

```{r}
head(christchurch)
```

## 3. Add month and year

```{r}
christchurch <- christchurch %>%
  mutate(month_year = "October 2025")
```

```{r}
count(christchurch, month_year)
```

## 4. Check missing values

```{r}
colSums(is.na(christchurch))
```

## 5. Check duplicates

```{r}
sum(duplicated(christchurch))
```

## 6. Check data types

```{r}
glimpse(christchurch)
```

## 7. Convert last_review to a date

```{r}
christchurch <- christchurch %>%
  mutate(last_review = as.Date(last_review))
```

## 8. Repeat for all 9 datasets

```{r}
data_dir <- "C:/Users/sohit/Downloads/UC - MS(ADS)/DATA 422 (Data Wrangling)/Project/Newzealand-Airbnb-Analysis/DATA"
files <- c(
  "October 2025" = "October_2025.csv",
  "November 2025" = "November_2025.csv",
  "December 2025" = "December_2025.csv",
  "January 2026" = "January_2026.csv",
  "February 2026" = "February_2026.csv",
  "March 2026" = "March_2026.csv",
  "April 2026" = "April_2026.csv",
  "May 2026" = "May_2026.csv",
  "June 2026" = "June 2026.csv"
)
files <- setNames(file.path(data_dir, files), names(files))
all_christchurch <- map2_dfr(
  files,
  names(files),
  ~ read_csv(.x, show_col_types = FALSE) %>%
    filter(neighbourhood_group == "Christchurch City") %>%
    mutate(
      month_year = .y,
      last_review = as.Date(last_review)
    )
)
```

## 9. Concatenate the datasets

The `map2_dfr()` step above reads and combines all nine datasets into one dataframe.

```{r}
christchurch_all <- all_christchurch

dim(christchurch_all)
```

```{r}
count(christchurch_all, month_year)
```

## 10. Check the final dataset

```{r}
head(christchurch_all)
```

```{r}
tail(christchurch_all)
```

```{r}
names(christchurch_all)
```

```{r}
dim(christchurch_all)
```

```{r}
count(christchurch_all, neighbourhood_group)
```

## 11. Check duplicates

```{r}
sum(duplicated(christchurch_all))
```

```{r}
sum(duplicated(christchurch_all$id))
```

## 12. Summary statistics for numerical columns

```{r}
numeric_summary <- christchurch_all %>%
  select(where(is.numeric)) %>%
  summarise(
    across(
      everything(),
      list(
        count = ~ sum(!is.na(.)),
        mean = ~ mean(., na.rm = TRUE),
        sd = ~ sd(., na.rm = TRUE),
        min = ~ min(., na.rm = TRUE),
        median = ~ median(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE)
      )
    )
  )

numeric_summary
```

## 13. Categorical variables and counts

```{r}
categorical_columns <- christchurch_all %>%
  select(where(~ is.character(.))) %>%
  names()

categorical_columns
```

```{r}
count(christchurch_all, room_type, sort = TRUE)
```

```{r}
count(christchurch_all, neighbourhood, sort = TRUE)
```

```{r}
count(christchurch_all, month_year, sort = FALSE)
```

## 14. General categorical summary

```{r}
categorical_summary <- map(
  categorical_columns,
  ~ christchurch_all %>%
    count(.data[[.x]], sort = TRUE) %>%
    slice_head(n = 20) %>%
    mutate(variable = .x)
)

categorical_summary
```

## 15. Missing-value summary

```{r}
missing_summary <- tibble(
  column = names(christchurch_all),
  missing_count = map_int(
    christchurch_all,
    ~ sum(is.na(.))
  ),
  missing_percentage = map_dbl(
    christchurch_all,
    ~ round(mean(is.na(.)) * 100, 2)
  )
)

missing_summary
```

## 16. Save the final concatenated dataset

```{r}
dir.create("output", showWarnings = FALSE, recursive = TRUE)

write_csv(
  christchurch_all,
  "output/Christchurch_listings_Oct2025_Jun2026.csv"
)
```

## 17. Save summary statistics

```{r}
write_csv(
  numeric_summary,
  "output/numerical_summary.csv"
)
```

```{r}
write_csv(
  missing_summary,
  "output/missing_values_summary.csv"
)
```

## 18. Check columns

```{r}
names(christchurch_all)
```

## 19. Recreate the combined dataset for analysis

```{r}
data_dir <- "C:/Users/sohit/Downloads/UC - MS(ADS)/DATA 422 (Data Wrangling)/Project/Newzealand-Airbnb-Analysis/DATA"
files <- file.path(data_dir, files)
files <- c(
  "October 2025" = "October_2025.csv",
  "November 2025" = "November_2025.csv",
  "December 2025" = "December_2025.csv",
  "January 2026" = "January_2026.csv",
  "February 2026" = "February_2026.csv",
  "March 2026" = "March_2026.csv",
  "April 2026" = "April_2026.csv",
  "May 2026" = "May_2026.csv",
  "June 2026" = "June 2026.csv"
)
files <- setNames(file.path(data_dir, files), names(files))

df <- map2_dfr(
  files,
  names(files),
  ~ read_csv(.x, show_col_types = FALSE) %>%
    filter(neighbourhood_group == "Christchurch City") %>%
    mutate(
      month_year = .y,
      month_year_date = case_when(
        TRUE ~ as.Date(paste0(.y, "-01"), format = "%B %Y-%d")
      )
    )
)
```

## 20. Check the combined data

```{r}
names(df)
```

```{r}
count(df, month_year)
```

## 21. Average Airbnb price by month

```{r}
month_order <- c(
  "October 2025",
  "November 2025",
  "December 2025",
  "January 2026",
  "February 2026",
  "March 2026",
  "April 2026",
  "May 2026",
  "June 2026"
)

df <- df %>%
  mutate(
    month_year = factor(
      month_year,
      levels = month_order,
      ordered = TRUE
    )
  )

monthly_price <- df %>%
  group_by(month_year) %>%
  summarise(
    average_price = mean(price, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(monthly_price, aes(x = month_year, y = average_price, group = 1)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Month",
    y = "Average Price (NZD)",
    title = "Average Airbnb Price in Christchurch: October 2025 – June 2026"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 22. Check month values

```{r}
head(df$month_year, 10)
```

## 23. Convert dates

The original Python workflow attempts to convert `month_year_date` to a date. In R, this column is created explicitly from `month_year`.

```{r}
df <- df %>%
  mutate(
    month_year_date = as.Date(
      paste0(
        as.character(month_year),
        " 01"
      ),
      format = "%B %Y %d"
    ),
    last_review = as.Date(last_review)
  )
```

```{r}
class(df$month_year_date)
class(df$last_review)
```

## 24. Calculate days since latest review

```{r}
df <- df %>%
  mutate(
    days_since_latest_review =
      as.numeric(month_year_date - last_review)
  )
```

```{r}
df %>%
  select(
    month_year_date,
    last_review,
    days_since_latest_review
  ) %>%
  head(20)
```

## 25. Histogram of days since latest review

```{r}
ggplot(
  df %>% filter(!is.na(days_since_latest_review)),
  aes(x = days_since_latest_review)
) +
  geom_histogram(bins = 50) +
  labs(
    x = "Days Since Latest Review",
    y = "Number of Listings",
    title = "Distribution of Days Since Latest Review"
  ) +
  theme_minimal()
```

## 26. Boxplot of days since latest review

```{r}
ggplot(
  df %>% filter(!is.na(days_since_latest_review)),
  aes(y = days_since_latest_review)
) +
  geom_boxplot() +
  labs(
    y = "Days Since Latest Review",
    title = "Boxplot of Days Since Latest Review"
  ) +
  theme_minimal()
```

## 27. Christchurch Airbnb price distribution

Only prices up to NZD 1,000 are included, matching the original Python analysis.

```{r}
df %>%
  filter(!is.na(price), price <= 1000) %>%
  ggplot(aes(x = price)) +
  geom_histogram(bins = 30) +
  labs(
    x = "Price (NZD)",
    y = "Frequency",
    title = "Christchurch Airbnb Price Distribution (Prices ≤ $1000)"
  ) +
  theme_minimal()
```

## 28. Publication date

The original Python code only assigns a publication date to October 2025. The same logic is reproduced here.

```{r}
df <- df %>%
  mutate(
    publication_date = case_when(
      as.character(month_year) == "October 2025" ~ as.Date("2025-10-05"),
      TRUE ~ as.Date(NA)
    )
  )
```

## 29. Recalculate days since latest review using publication date

```{r}
df <- df %>%
  mutate(
    days_since_latest_review =
      as.numeric(publication_date - last_review)
  )
```

```{r}
df %>%
  select(
    publication_date,
    last_review,
    days_since_latest_review
  ) %>%
  head(20)
```

## 30. Average days since latest review

```{r}
average_days <- mean(
  df$days_since_latest_review,
  na.rm = TRUE
)

print(
  paste(
    "Average days since latest review:",
    round(average_days, 2)
  )
)
```

## 31. Histogram of review age

```{r}
review_days <- df$days_since_latest_review %>%
  na.omit()

bins <- c(
  0, 50, 100, 150, 200, 250, 300, 350,
  400, 450, 500, 550, 600, 650, 700,
  750, 800, 850, 900, 950, 1000
)

ggplot(
  tibble(review_days = review_days),
  aes(x = review_days)
) +
  geom_histogram(
    breaks = bins
  ) +
  scale_x_continuous(breaks = bins) +
  labs(
    x = "Days Since Latest Review",
    y = "Number of Properties",
    title = "Days Since Latest Review - Christchurch Airbnb Properties (October 2025)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
