---
title: "Deliverable 5"
author: "Anibal"
date: "2026-09-17"
output: html_document
---
```{r}
library(tidyverse)
library(jsonlite)
library(httr)
library(parallel)
```


```{r}
christchurch_clean <- read_csv(
  "data/clean/christchurch_clean.csv",
  show_col_types = FALSE
)

tenancy_clean <- read_csv(
  "data/clean/tenancy_clean.csv",
  show_col_types = FALSE
)
```

```{r}
names(christchurch_clean)
```
```{r}
names(tenancy_clean)
```

```{r}
christchurch_clean %>%
  select(latitude, longitude) %>%
  drop_na() %>%
  slice(1)
```

```{r}
.api_key <- '97cfd31039bc471789e523737d341432'
```


```{r}

coords <- christchurch_clean %>%
  select(latitude, longitude) %>%
  drop_na() 
```
```{r}
nrow(coords)
```

```{r}
get_area_code_parallel <- function(coords) {
  
  url <- paste0(
    "https://koordinates.com/services/query/v1/vector.json",
    "?key=", .api_key,
    "&layer=123515",
    "&x=", coords$longitude,
    "&y=", coords$latitude,
    "&radius=10000"
  )
  
  for (i in 1:3) {
    
    result <- tryCatch({
      
      response <- fromJSON(url, simplifyVector = FALSE)
      
      features <- response$vectorQuery$layers[["123515"]]$features
      
      if (length(features) == 0) {
        return(NA_character_)
      }
      
      features[[1]]$properties$SA22026_V1_00
      
    }, error = function(e) {
      NULL
    })
    
    if (!is.null(result)) {
      return(result)
    }
    
    Sys.sleep(1)
  }
  
  NA_character_
}
```

```{r}
cl <- makeCluster(8)

clusterExport(cl, c("get_area_code_parallel", ".api_key"))
clusterEvalQ(cl, library(jsonlite))

area_codes <- parLapply(
  cl,
  split(coords, seq_len(nrow(coords))),
  get_area_code_parallel
)

stopCluster(cl)

area_codes <- unlist(area_codes)
```
```{r}
length(area_codes)
```
```{r}
sum(is.na(area_codes))
```
```{r}
head(area_codes)
```
```{r}
christchurch_clean$area_code <- area_codes
```

```{r}
christchurch_clean %>%
  select(latitude, longitude, area_code) %>%
  slice(1:10)
```
```{r}
write_csv(
  christchurch_clean,
  "data/clean/christchurch_clean_area_codes.csv"
)
```

```{r}
file.exists("data/clean/christchurch_clean_area_codes.csv")
```

```{r}
names(tenancy_clean)
```

```{r}
head(tenancy_clean$`Location Id`)
```
```{r}
head(christchurch_clean$month_year)
```

```{r}
head(tenancy_clean$TimeFrame)
```
```{r}
tenancy_clean <- tenancy_clean %>%
  mutate(
    month_year = format(as.Date(TimeFrame), "%B_%Y")
  )
```

```{r}
head(tenancy_clean$month_year)
```

```{r}
tenancy_clean <- tenancy_clean %>%
  mutate(area_code = as.character(`Location Id`))
```


```{r}
tenancy_clean %>%
  count(area_code, month_year) %>%
  filter(n > 1) %>%
  slice_head(n = 10)
```
```{r}
tenancy_clean %>%
  filter(
    area_code == "100100",
    month_year == "April_2026"
  ) %>%
  select(
    area_code,
    month_year,
    `Dwelling Type`,
    `Number Of Beds`,
    `Median Rent`
  )
```
```{r}
tenancy_join <- tenancy_clean %>%
  filter(
    `Dwelling Type` == "ALL",
    `Number Of Beds` == "ALL"
  )
```

```{r}
tenancy_join %>%
  count(area_code, month_year) %>%
  filter(n > 1)
```

```{r}
christchurch_joined <- christchurch_clean %>%
  left_join(
    tenancy_join,
    by = c("area_code", "month_year")
  )
```

```{r}
nrow(christchurch_joined)
```
```{r}
christchurch_joined %>%
  filter(area_code == "326600") %>%
  summarise(
    median_airbnb_price = median(price, na.rm = TRUE),
    number_of_airbnbs = n()
  )
```

Median Price for an AirBnB in Christchurch Central (326600) is $240 and there are 715 AirBnbs in Christchurch Central.

```{r}
christchurch_joined <- christchurch_joined %>%
  mutate(
    long_term_price_per_night = `Median Rent` / 7,
    price_difference = price - long_term_price_per_night
  )
```

```{r}
christchurch_joined %>%
  group_by(area_code) %>%
  summarise(
    median_gap = median(price_difference, na.rm = TRUE),
    number_of_airbnbs = n()
  ) %>%
  arrange(desc(median_gap)) %>%
  slice_head(n = 10)
```

```{r}
christchurch_joined %>%
  filter(area_code %in% c("322800", "328800", "332700", "322300", "317400")) %>%
  group_by(area_code) %>%
  summarise(
    median_gap = median(price_difference, na.rm = TRUE),
    number_of_airbnbs = n()
  )
```
```{r}
christchurch_joined %>%
  filter(area_code %in% c("317400", "322300", "322800", "328800", "332700")) %>%
  group_by(area_code) %>%
  summarise(
    median_gap = median(price_difference, na.rm = TRUE),
    number_of_airbnbs = n()
  ) %>%
  arrange(desc(median_gap))
```

```{r}
christchurch_joined %>%
  filter(area_code %in% c("317400", "322300", "322800", "328800", "332700")) %>%
  select(area_code, neighbourhood) %>%
  distinct() %>%
  arrange(area_code)
```

```{r}
area_names <- christchurch_joined %>%
  filter(area_code %in% c("317400", "322300", "322800", "328800", "332700")) %>%
  group_by(area_code) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    area_name = map2_chr(
      latitude,
      longitude,
      function(lat, lon) {
        url <- paste0(
          "https://koordinates.com/services/query/v1/vector.json",
          "?key=", .api_key,
          "&layer=123515",
          "&x=", lon,
          "&y=", lat,
          "&radius=10000"
        )
        
        response <- fromJSON(url, simplifyVector = FALSE)
        
        response$vectorQuery$layers[["123515"]]$features[[1]]$properties$SA22026_V1_00_NAME
      }
    )
  ) %>%
  select(area_code, area_name) %>%
  distinct()
```

```{r}
area_names
```
```{r, fig.width= 15, fig.height = 12}
top_gap_areas <- christchurch_joined %>%
  filter(!is.na(price_difference)) %>%
  group_by(area_code) %>%
  summarise(
    median_gap = median(price_difference),
    number_of_airbnbs = n()
  ) %>%
  filter(number_of_airbnbs >= 10) %>%
  arrange(desc(median_gap)) %>%
  slice_head(n = 10)
```

```{r}
christchurch_joined %>%
  filter(area_code %in% top_gap_areas$area_code) %>%
  ggplot(aes(x = factor(area_code), y = price_difference)) +
  geom_boxplot() +
  labs(
    title = "Airbnb versus long-term rental price gaps",
    x = "Area code",
    y = "Price difference per night ($)"
  )
```
```{r}
top_gap_areas
```
The largest typical gap between short-term Airbnb and long-term rental prices was observed in Sumner (SA2 332700), where the median difference was approximately $197.43 per night, based on AirBnB prices minus the median long-term weekly rent converted to a nightly rate.

```{r}
tenancy_join %>%
  select(
    area_code,
    month_year,
    `Total Bonds`
  ) %>%
  head(10)
```

```{r}
location_comparison <- christchurch_joined %>%
  filter(month_year == "April_2026") %>%
  group_by(area_code) %>%
  summarise(
    airbnb_count = n(),
    rental_properties = first(`Total Bonds`)
  )
```

```{r}
location_comparison %>%
  arrange(desc(airbnb_count)) %>%
  head(10)
```

```{r}
location_comparison_clean <- location_comparison %>%
  filter(!is.na(rental_properties))
```

```{r}
location_comparison_clean %>%
  arrange(desc(airbnb_count)) %>%
  head(10)
```

```{r}
location_comparison_clean %>%
  arrange(desc(airbnb_count)) %>%
  slice_head(n = 10) %>%
  pivot_longer(
    cols = c(airbnb_count, rental_properties),
    names_to = "property_type",
    values_to = "count"
  ) %>%
  ggplot(aes(x = reorder(area_code, count), y = count, fill = property_type)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    title = "Airbnb and Rental Properties by Area",
    x = "Area code",
    y = "Number of properties",
    fill = "Property type"
  )
```
The key comparison is visible:

327000: 318 Airbnb vs 141 rentals
327100: 246 vs 42
325800: 168 vs 105
326600: 138 vs 6
325700: 126 vs 54


```{r}
file.exists("data/clean/christchurch_clean_area_codes.csv")
```






